
```{r}
library(rlang)
library(lobstr)
```

Outline: 

- 17.2 : Code is data. This teaches you how to create a modify expressions by capturing code 
- 17.3 : Describes the tree-like structure of code
- 17.4 : Shows how to create new expressions programmatically 
- 17.5 : Shows how to execute expressions by evaluating them in an environment 
- 17.6 : Illustrates how to customize evaliation by supplying custom functions in a new environment 
- 17.7 : extends that customization into data masks, which blur the line between environments and dataframes 
- 17.8 : Introduces a new data structure called the quosure

## 17.2 | Code is Data 

We can capture code and compute on it as we would any other type of data. 

```{r}
# expr() essentially returns exactly what we pass in
expr(mean(x, na.rm = TRUE))
expr(10 + 100 + 1000)
```

Captured code is called an **expression**. This isn't a single type of object, but rather a collective term for any of four types: call, symbol, constant, or pairlist. 


expr() lets us capture code that we've typed - but we need a different tool for use inside of functions

```{r}
capture_it <- function(x) {
    expr(x)
}

capture_it(a + b + c)

# enexpr() or enrich expr() takes a lazily evaluated argument and turns it into an expr
capture_it <- function(x) {
    enexpr(x)
}

capture_it(a + b + c)
```

Since capture_it() uses enexpr(), we say that it automatically quotes its first argument. Once we have captured an expression, we can inspect and modify it (using [[]] and \$)

```{r}
f <- expr(f(x = 1, y = 2))

# add a new argument
f$z <- 3

f

# or remove an argument
f[[2]] <- NULL

f
```

The first element of the call is the function to be called, which means that the first argument is in the second position. 

## 17.3 | Code is a Tree

Behind the scenes, almost every programming language represents code as a tree, often called an **abstract syntax tree**. In R we can actually inspect and manipulate this tree.

```{r}
ast(f(a, "b"))

ast(f1(f2(a, b), f3(1, f4(2))))
```

Since all function forms can be written in prefix form, every R expression can be displayed in this way

```{r}
ast(1 + 2 * 3)
```

## 17.4 | Code Can Generate Code 

As well as seeing the tree from code typed by a human, we can also use code to create new trees. There are two main tools: call2() and unquoting 

call2 constructs a function call from its components: the function to call and the arguments to call it with. 

```{r}
call2("f", 1, 2, 3)
call2("+", 1, call2("*", 2, 3))
```

call2 is often convenient to program with, but is a bit clunky for interactive use. An alternative technique is to build complex code trees by combining simpler code trees with a template. expr() and enexpr() have built in support for this idea via !!, the unquote operator. 

Basically, !!x inserts a code tree stored in x into the expression. This makes it easy to build complex trees from simple fragments. 

```{r}
xx <- expr(x + x)
yy <- expr(y + y)

expr(!!xx / !!yy)
```

Notice that the output preserves order. 

Unquoting gets more useful when you wrap it into a function, first using enexpr() to capture the user's expression, then expr() and !! to create a new expression using a template. 

```{r}
# generate an expression that computes the coefficient of variation
cv <- function(var) {
    var <- enexpr(var)
    expr(sd(!!var)  / mean(!!var))
}

cv(x)

cv(x + y)
```

This isn't very useful here, but being able to create this sort of building block is very useful when solving more complex problems. Importantly, this works even when given weird variable names 

```{r}
cv(`(`)
```

Dealing with weird names is another reason to avoid paste() when generating R code. This may seem an esoteric concern, but not worrying about it when generating SQL code in web applications lead to SQL injection attacks that have collectively cost billions of dollars. 

## 17.5 | Evaluation Runs Code 

Inspecting and modifying code gives us one set of powerful tools - another set is provided when we evaluate (execute or run) an expression. 

The primary tool for evaluating expressions is base::eval(), which takes an expression and an environment 

```{r}
eval(expr(x + y), env(x = 1, y = 10))
eval(expr(x + y), env(x = 2, y = 100))

# if we omit the env var, eval uses the current env
x <- 10
y <- 100
eval(expr(x + y))
```

A big advantage of evaluating code manually is that we can tweak the environment. There are two main reasons for this: 

- To temporarily override functions to implement a domain specific language 
- To add a data mask so we can refer to variables in a dataframe as if they were variables in an environment 

## 17.6 | Customizing Evaluation with Functions

In addition to using an environment that can bind x and y to vectors, we can also bind names to functions, allowing us to override the behaviour of existing functions. 

Here is an example 

```{r}
string_math <- function(x) {
    e <- env(
        caller_env(),
        `+` = function(x, y) paste0(x, y),
        `*` = function(x, y) strrep(x, y)
    )
    eval(enexpr(x), e)
}

name <- "Michael"
string_math("Hello " + name)

string_math(("x" * 2 + "-y") * 3)
```

dplyr takes this idea to the extreme, running code in an environment that generates SQL for execution in a remote database 

```{r}
library(dplyr)

con <- DBI::dbConnect(RSQLite::SQLite(), filename = ":memory:")

mtcars_db <- copy_to(con, mtcars)

mtcars_db %>%
    filter(cyl > 2) %>%
    select(mpg:hp) %>%
    head(10) %>%
    show_query()

DBI::dbDisconnect(con)
```

## 17.7 | Customizing Evaluation with Data

Rebinding functions is a powerful technique, but it requires some time investment when considering how it will be used. A more immediately practical application is modifying evaluation to look for varaibles in a data frame instead of an environment. 

```{r}
df <- data.frame(x = 1:5, y = sample(5))

eval_tidy(expr(x + y), df)
```

Evaluating with a data mask is a useful technique for interactive analysis because it allows you to write x + y instead of df$x + df$y. 

We can wrap this pattern up into a function using enexpr() giving us something similar to base::with()

```{r}
with2 <- function(df, expr) {
    eval_tidy(enexpr(expr), df)
}

with2(df, x + y)
```

Unfortunately, this function has a subtle bug which will require a new data structure to help deal with it. 

## 17.8 | Quosures 



To see the problem, we can modify with2

```{r}
with2 <- function(df, expr) {
    a <- 1000
    eval_tidy(enexpr(expr), df)
}
```

We can see the problem wehn we use with2 to refer to a variable caled a. We want the value of a to come from the binding we can see (10), not the binding internal to the function (1000)

```{r}
df <- data.frame(x = 1:3)
a <- 10
with2(df, x + a)
```

This arises because we need to evaluate the captured expression in the environment where it was written (where a is 10), not the environment inside of with2() where a is 1000. 

We can solve this problem with a new data structure: the **quosure**. The quosure bundles an expression with the environment. eval_tidy() knows how to work with quosures, so all we need to do is switch out enexpr() for enquo()

```{r}
with2 <- function(df, expr) {
    a <- 1000
    eval_tidy(enquo(expr), df)
}

with2(df, x + a)
```

Whenever we use a data mask, we must always use enquo instead of enexpr. 

