# Expressions 

# 18.1 | Introduction

```{r}
library(tidyverse)
library(lobstr)
library(rlang)
```

To compute on the language, we first need to understand the structure of the language. This requires some new vocabulary, tools and ways of thinking.

The first of these is the distinction between an operation and its result. 

Take the following: it multiplies a variable x by 10 and saves the result to a new variable called y. It doesn't work because we haven't defined a variable x

```{r}
y <- x * 10
```

It would be nice if we could capture the intent of the code without executing it. In other words, we wish to separate our description of the action from the action itself. 

```{r}
(z <- rlang::expr(y <- x * 10))
```

expr() returns an expression, an object that captures the structure of the code without evaluating it. 

If you have an expression, you can evaluate it with base::eval

```{r}
x <- 4

eval(z)

y
```

The focus of this cahpter is the data structures that underlie expressions. Mastering this will allow us to inspect and modify captured code, and to generate code with code. 

## Outline 

- 18.2 introduces the idea of an abstract syntax tree
- 18.3 dives into the data structures that underpin the AST: constants, symbols, and calls, which are collectively known as expressions
- 18.4 covers parsing, the act of converting the linear sequence of character in code into the AST
- 18.5 Shows how to use recursive functions to compute on the language, writing functions that compute with expressions 
- 18.6 Looks at three more specialized data structures: pairlists, missing arguments, and expression vectors 

# 18.2 | Abstract Syntax Trees 

Expressions are also called abstract syntax trees because the structure of code is hierarchical and can be naturally represented as a tree. Understanding this tree structure is crucial for inspecting and modifying expressions (i.e. metaprogramming)

## 18.2.1 | Drawing 

```{r}
ast(f(x, "y", 1))
```

```{r}
ast(f(g(1, 2), h(3, 4, i())))
```

These can be read top to bottom. This also generally determines evaluation order, as evaluation generally proceeds from the deepest to the shallowest, but this is not guaranteed due to lazy evaluation. 

## 18.2.2 | Non-Code Components 

These are _abstract_ syntax trees because they only capture important structural details of the code, not whitespace or comments 

```{r}
ast(
    f(x, y) # important comment 
)

# theres only one place where whitespace affects the AST
ast(y <- x)
ast(y < -x)
```

## 18.2.3 | Infix Calls 

Every call in R can be written in tree form because any call can be written in prefix form. 

```{r}
# these have the same AST
ast(y <- x * 10)

ast(`<-`(y, `*`(x, 10)))

# if you generate an expression with prefix calls, R will still print it in infix form
expr(`<-`(y, `*`(x, 10)))
```

## 18.2.4 | Exercises 

1. 

```{r}
ast(f(g(h())))

ast(1 + 2 + 3)

ast(z * (x + y))
```

3. What is happening with the ASTs below? 

```{r}
ast(`x` + `y`)

ast(x ** y)

ast(1 -> x)

ast(if (x < 5) {x} else if (x < 6) {x + 1} else if (x < 7) {x + 2} else if (x < 8) {y})
```

# 18.3 | Expressions 

An expression is any member of the set of base types created by parsing code: 

- constant scalars 
- symbols 
- call objects 
- pairlists 

These are the data structures used to represent captured code from expr() and is_expresson(expr(...)) is always true. 

## 18.3.1 | Constants 

A constant is either NULL, or a length 1 atomic vector like TRUE, 1L, 2.5 or "x". We can test for a constant with rlang::is_syntactic_literal()

Constants are self quoting in the sense that the expression used to represent a constant is the same constant. 

```{r}
identical(expr(TRUE), TRUE)
```

## 18.3.2 | Symbols 

A symbol represents the name of an object, like x, mtcars, or mean. We can create a symbol in two ways: by capturing code that references an object with expr(), or turning a string into a symbol with rlang::sym()

```{r}
expr(x)

sym("x")

# turn a symbol back into a string
as_string(expr(x))

# tells if its a symbol
str(expr(x))

is.symbol(expr(x))
```

The symbol type is not vectorized, i.e. a symbol is always length 1. If we want multiple symbols, we must put them in a list (using something like syms())

## 18.3.3 | Calls 

A call object represents a captured function call. Call objects are a special type of list, where the first component specifies the function to call (usually a symbol), and the reamining elements are the arguments for that call. 

Confusingly, typeof() and str() print "language" for calls, but is.call() returns TRUE

```{r}
ast(read.table("important.csv", row.names = FALSE))

x <- expr(read.table("important.csv", row.names = FALSE))

typeof(x)
is.call(x)
```

### 18.3.3.1 | Subsetting 

Calls generally behave like lists, i.e. we have to use standard subsetting tools. The first element of the call object is the function to call (usually a symbol). 

```{r}
x[[1]]

is.symbol(x[[1]])

# the remainded to the elements are the arguments
as.list(x[-1])

x$row.names

# determine the number of arguments
length(x) - 1
```

Extracting specific arguments from calls is challenging because of R's flexible rules for argument matching: it could potentially be in any location, with the full name, an abbreviated name, or no name at all. 

To alleviate this, we can use call_standardise() which standardizes all arguments to use the full name

```{r}
call_standardise(x)

x$header <- TRUE

x
```

### 18.3.3.2 | Function Position 

the first element of the call object is the function position. This contains the function that will be called when the object is evaluated, and is usually a symbol

```{r}
ast(foo())

# when we use quotes, the parser converts it to a symbol
ast("foo"())
```

If the function doesn't exist in the current environment, the function position will be occupied by another call

```{r}
ast(pkg::foo(1))
```

### 18.3.3.3 | Constructing 

We can construct a call object from its components using call2(). The first argument is the name of the function to call (either as a string, a symbol, or another call). The remaining args are passed to the call 

```{r}
call2("mean", x = expr(x), na.rm = TRUE)

call2(expr(base::mean), x = expr(x), na.rm = TRUE)

# infix calls still print as usual
call2(`<-`, expr(x), 10)
```

# 18.4 | Parsing and Grammar 

The process by which a computer language takes a string and constructs an expression is called parsing, and it is governed by a set of rules known as a grammar. 

In this section we use ast() to explore some of the details of R's grammar, and show how you can transform between expressions and strings.

## 18.4.1 | Operator Precedence 

Infix operators introduce two sources of ambiguity: 

1. What does 1 + 2 * 3 yield? 

Programming languages use conventions called operator precedence to resolve this ambiguity. 

```{r}
ast(1 + 2 * 3)
```

Arithmetic operators are straight forward because they are generally the same across languages.

Predicting the precedence of other operators is harder. One particularly surprising case is the ! operator, which has a much lower precedence than expected.

```{r}
ast(!x %in% y)
```

R has over 30 infix operators divided into 18 precedence groups. Details are described in ?Syntax

Details:

The following unary and binary operators are defined.  They are
listed in precedence groups, from highest to lowest.
    
‘:: :::’           access variables in a namespace              
‘$ @’              component / slot extraction                  
‘[ [[’             indexing                                     
^                exponentiation (right to left)               
‘- +’              unary minus and plus                         
:                sequence operator                            
‘%any%’ special operators (including %% and %/%)
‘* /’              multiply, divide                             
‘+ -’              (binary) add, subtract                       
‘< > <= >= == !=’  ordering and comparison                      
!                negation                                     
‘&  &&’            and                                          
‘| ||’             or                                           
~                as in formulae                               
‘-> ->>’           rightwards assignment                        
‘<- <<-’           assignment (right to left)                   
=                assignment (right to left)                   
?                help (unary and binary)                      

## 18.4.2 | Associativity

The second source of ambiguity is introduced by repeated usage of the same infix function. 

In R most operators are left-associative, i.e. the operations on the left are evaluated first

```{r}
ast(1 + 2 + 3)
```

There are two exceptions: exponentiation and assignment

```{r}
ast(2^2^3)

ast(x <- y <- z)
```

## 18.4.3 | Parsing and Deparsing 

Most of the time when we type code into the console R takes care of turning the characters into an AST. Occasionally we have code stored in a string and we want to parse it ourselves. We can do so with parse_expr()

```{r}
x1 <- "y <- x + 10"

x1

is.call(x1)

(x2 <- parse_expr(x1))

is.call(x2)
```

parse_expr always returns a single expression. If we have multiple expressions separated by ; or \n we need to use parse_exprs(). 

```{r}
x3 <- "a <- 1; a + 1"

parse_exprs(x3)
```

If we find ourselves working with strings containing code frequently, we should reconsider the process; read ch 19 and consider whether we can generate expressions using quasiquotation more safely.

The inverse of parsing is deparsing. Given an expression, we want the string that would generate it. 

This happens automatically when we print an expression.

```{r}
z <- expr(y <- x + 10)

expr_text(z)
```

Parsing and deparsing are not perfectly symmetric because parsing generates an abstract syntax tree. This means we lose backticks around ordinary names, comments, and whitespace. 

```{r}
cat(expr_text(expr({
    # comment here
    x <-    `x` + 1
})))
```

## 18.4.4 | Exercises 

1.

```{r}
ast(f((1)))

ast(`(`(1 + 1))
```

3.

```{r}
-2^2

ast(-2^2)

(-2)^2

ast((-2)^2)
```

4.

```{r}
!!1 + !!1
```

5. 

```{r}
x1 <- x2 <- x3 <- x4 <- 0

ast(x1 <- x2 <- x3 <- x4 <- 0)
```

6. 

```{r}
ast(x + y %+% z)

ast(x ^ y %+% z)
```

7. 

```{r}
parse_exprs("x + 1; y + 1")
```

8. 

```{r}
parse_expr("f()")
```

9. 

```{r}
expr <- expr(g(a + b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s))

deparse(expr)

expr_text(expr)
```

```{r}
pairwise.t.test
```

# 18.5 | Walking AST with Recursive Functions 

This section uses everything learned about ASTs to solve a more complicated problem. The inspiration comes from the base codetools package, which provides two interesting functions:

1. findGlobals() : locates all global variables used by a function. 
2. checkUsage() : checks for a range of common problems including unused local variables, parameters, and use of partial argument matching

We will focus on the big underlying idea: recursion on the AST.

Recursive functions have two parts:

The recursive case and the base case. To illuminate the pattern, we need two helper functions

```{r}
# return name of expression type
expr_type <- function(x) {
    if (rlang::is_syntactic_literal(x)) {
        "constant"
    } else if (is.symbol(x)) {
        "symbol"
    } else if (is.call(x)) {
        "call"
    } else if (is.pairlist(x)) {
        "pairlist"
    }
     else {
        typeof(x)
    }
}

check_these <- c(expr("a"), expr(x), expr(f(1, 2)))

check_these %>% map(expr_type)

# wrapper around the switch function
switch_expr <- function(x, ...) {
    switch(expr_type(x),
           ...,
           stop("Don't know how to handle type ", typeof(x), call. = FALSE))
}

```

With these two functions , we can write a basic template for any function that walks the AST using switch.

```{r}
recurse_call <- function(x) {
    switch_expr(x,
                # base cases
                symbol = ,
                constant = ,
                # recursive cases
                call = ,
                pairlist = )
}
```

Typically, solving the base case is easy, and the recursive cases are easier. 

### 18.5.1 | Finding F and T

```{r}
# return TRUE if input uses logical abbreviation (i.e., T = TRUE, F = FALSE)

# first find types for each
expr_type(expr(TRUE))
expr_type(expr(T))

# TRUE is parsed as a logical vector of length 1
# T is parsed as a name

logical_abbr_rec <- function(x) {
    switch_expr(x,
                constant = FALSE,
                symbol = as_string(x) %in% c("F", "T"))
}

try_these <- c(expr(T), expr(TRUE), expr(F), expr(FALSE))

try_these %>% map(logical_abbr_rec)
```

We've written logical_abbr_rec assuming that the input will be an expression as this will make the recursive operation simpler. When writing a recursive function its common to write a wrapper that provides defaults or makes the function easier to use. Typically we make a wrapper that quotes its input so we don't need to use expr() every time.

```{r}
logical_abbr <- function(x) {
    logical_abbr_rec(enexpr(x))
}
```

Next we must implement the recursive cases. Here we want to do the same thing for calls and pairlists: recursively apply the function to each subcomponent and return true if any subcomponents contains a logical abbreviation.

```{r}
logical_abbr_rec <- function(x) {
    switch_expr(x,
                constant = FALSE,
                symbol = as_string(x) %in% c("T", "F"),
                call = ,
                pairlist = some(x, logical_abbr_rec))
}

logical_abbr(mean(x, na.rm = T))
logical_abbr(mean(x, na.rm = TRUE))
logical_abbr(function(x, na.rm = T) FALSE)
```

### 18.5.2 | Finding all variables created by assignment 

The next task is to list all the variables created by assignment. This is a little more complicated. 

```{r}
# start with AST for assignment
ast(x <- 10)
```

Assignment is a call object where the first element is the symbol <- , the second is the name of the variable, and the third is the value to be assigned. 

Next we decide which data structure to use for the results. 

```{r}
# implement the base case
find_assign_rec <- function(x) {
    switch_expr(x,
                constant = ,
                symbol = character())
}

find_assign <- function(x) find_assign_rec(enexpr(x))

find_assign(x)
find_assign("x")
```

Next we implement the recursive cases. This is made easier by a function that should exist in purrr, but doesn't (currently).

```{r}
# expects a .f to return a character vector of arbitrary length and flattens all results into a single character vector 
flat_map_chr <- function(.x, .f, ...) {
    flatten_chr(map(.x, .f, ...))
}

flat_map_chr(letters[1:3], ~ rep(., sample(3, 1)))
```

The recursive case for pairlists is straight forward: We iterate over every element of the pairlist (i.e. each function argument) and combine the results. The case for calls is a little more complex : if this is a call to <- then we should return the second element of the call

```{r}
find_assign_rec <- function(x) {
    switch_expr(x,
                # base cases
                constant = ,
                symbol = character(),
                # recursive
                pairlist = flat_map_chr(as.list(x), find_assign_rec),
                call = {
                    if (is_call(x, "<-")) {
                        as_string(x[[2]])
                    } else {
                        flat_map_chr(as.list(x), find_assign_rec)
                    }
                })
}

find_assign(a <- 1)

find_assign({a <- 1
    {
        b <- 2
    }
})
```

Now we need to make our functions more robust by finding examples intended to break it. 

What happens wehn we assign to the same variable multiple times?

```{r}
find_assign({
    a <- 1
    a <- 2
})

# fix as the level of the wrapper function
find_assign <- function(x) unique(find_assign_rec(enexpr(x)))

find_assign({
    a <- 1
    a <- 2
})
```

What happens if we make nested calls to <- ? Currently we only return the first, because when <- appears we terminate recursion.

```{r}
find_assign({
    a <- b <- c <- 1
})
```

Instead we need a more rigorous approach. Its best to keep the recursive function focused on the tree structure, so we extract out find_assign_call into a separate function

```{r}
find_assign_call <- function(x) {
    if (is_call(x, "<-") && is_symbol(x[[2]])) {
        lhs <- as_string(x[[2]])
        children <- as.list(x)[-1]
    } else {
        lhs <- character()
        children <- as.list(x)
    }
    c(lhs, flat_map_chr(children, find_assign_rec))
}

find_assign_rec <- function(x) {
    switch_expr(x,
                # base cases
                constant = ,
                symbol = character(),
                # recursive cases
                pairlist = flat_map_chr(x, find_assign_rec),
                call = find_assign_call(x))
}

find_assign(a <- b <- c <- 1)

find_assign(system.time(x <- print(y <- 5)))
```

### 18.5.3 | Exercises 

1. 

```{r}
logical_abbr_rec

logical_abbr(T(1, 2, 3))

is_call(expr(T(1, 2, 3)))

to_pass <- expr(T(1, 2, 3))

is_call(to_pass)

is_call(expr("1, 2, 3"))

logical_abbr_rec <- function(x) {
    switch_expr(x,
                constant = FALSE,
                symbol = if (is.call(x)) {
                             FALSE
                         } else {
                             as_string(x) %in% c("T", "F")
                         },
                call = ,
                pairlist = some(x, logical_abbr_rec))
}

```

## 18.6 | Specialized Data Structures

There are two data structures and one special symbol which will be covered for the sake of completeness. They are usually not important in practice.


### 18.6.1 | Pairlists 

pairlists are a remnant of R's past and have been replaced by lists almost everywhere. The only place we are likely to see pairlists in R is when working with ccalls to the function function, as the formal arguments to a function are stored in a pairlist.

```{r}
f <- expr(function(x, y = 10) x + y)

args <- f[[2]]

args

typeof(args)
```

Fortunately, whenever we encounter a pairlist, we can treat it just like a regular list. 

```{r}
pl <- pairlist(x = 1, y = 2)

pl %>% length()

pl$x 
```

Behind the scenes, pairlists are implemented using a linked list rather than an array. 

### 18.6.2 | Missing Arguments 

The special symbol that needs a little extra discussion is the empty symbol, which is used to represent missing arguments (not missing values !). We only need to care about the missing symbol if we are programmatically creating functions with missing arguments.

We can make one with missing_arg() or expr()

```{r}
missing_arg()

typeof(missing_arg())

# check if we have one since it doesn't print anything
is_missing(missing_arg())
```

We can find them in the wild in function formats:

```{r}
f <- expr(function(x, y = 10) x + y)

args <- f[[2]]

args[[1]] %>% is_missing()
```

This is particularly important for ..., which is always associated with an empty symbol:

```{r}
f <- expr(function(...) list(...))

args <- f[[2]]

is_missing(args[[1]])
```

If you bind an empty symbol to a variable, then access the variable, you will get an error

```{r}
(m <- missing_arg())

m
```

But you won't if you store it inside another data structure

```{r}
ms <- list(missing_arg(), missing_arg())

ms[1]
```

If we need to preserve the missingness of a variable, rlang::maybe_missing() is often helpful. It allows us to refer to a potentially missing variable without triggering the error.

### 18.6.3 | Expression Vectors 

Expression vectors are produced by only two base functions: expression() and parse()

```{r}
exp1 <- parse(text = c("
x <- 4
x
"))

exp2 <- expression(x <- 4, x)

exp1
exp2
typeof(exp1)
typeof(exp2)
```

Like calls and pairlists, expression vectors behave like lists. 

```{r}
length(exp1)

exp1[[1]]
```

Conceptually, an expression vector is just a list of expressions. The only difference is that calling eval() on an expression evaluates each individual expression. Generally we would just use lists of expressions.


