# Rewriting R code in C++

```{r}
library(Rcpp)
```

Typical bottlenecks that C++ can address include:

- Loops that can't be easily vectorized because subsequent iterations depend on previous ones 
- Recursive functions, or problems which involve calling functions millions of times. The overhead of calling a function is much lower in C++ than in R
- Problems that require advanced data structures and algorithms that R doesn't provide. 

The focus here is on reducing bottlenecks in R code, and creating small self contained functions. 

# Outline 

- 25.2 | teaches how to write some C++ 
- 25.3 | Modify attributes from C++ 
- 25.4 | Work with R's missing values in C++
- 25.5 | Use data structures from the standard template library, or STL, built into C++
- 25.6 | Two case studies 
- 25.7 | Add C++ code to a package 
- 25.8 | more resources 


# 25.2 | Getting Started 

```{r}
cppFunction('int add(int x, int y, int z) {
int sum = x + y + z;
return sum;
}')

add

add(1, 2, 3)
```

When ran, Rcpp will compile the C++ code and construct an R function that connects to the compiled C++ function. 

## 25.2.1 | No inputs, Scalar Output 

```{r}
# a very simple function
one <- function() 1L

cppFunction('int one() {
return 1;
}')

one()
```


## 25.2.2 | Scalar Input, Scalar Output

```{r}
signR <- function(x) {
    if (x > 0) {
        1
    } else if (x == 0) {
        0
    } else {
        -1
    }
}

cppFunction('int signC(int x) {
if (x > 0) {
return 1;
} else if (x == 0) {
return 0;
} else {
return -1;
}
}')

signC(-0)

# the cost of loops is much lower in C++
sumR <- function(x) {
    total <- 0
    for (i in seq_along(x)) {
        total <- total + x[i]
    }
    total
}

cppFunction('double sumC(NumericVector x) {
    int n = x.size();
    double total = 0;
    for (int i = 0; i < n; ++i) {
        total += x[i];
    }
    return total;
}')

x <- runif(1e3)

bench::mark(
           sum(x),
           sumC(x),
           sumR(x)
)[1:6]
```

## 25.2.4 | Vector Input, Vector Output 

```{r}
# create a function that computes Euclidean distance between a value and a vector of values
pdistR <- function(x, ys) {
    sqrt((x - ys) ^ 2)
}

cppFunction('NumericVector pdistC(double x, NumericVector ys) {
    int n = ys.size();
    NumericVector out(n);

    for (int i = 0; i < n; ++i) {
        out[i] = sqrt(pow(ys[i] - x, 2.0));
    }
    return out;
}')

# because the R version is fully vectorized, it will already be fast
y <- runif(1e6)

bench::mark(
           pdistR(0.5, y),
           pdistC(0.5, y)
)[1:6]
```

## 25.2.5 | Using sourceCpp

you can read in Cpp source code with sourceCpp("path/to/file.cpp"). 

## 25.2.6 | Exercises 

```{r}
cppFunction('double f1(NumericVector x) {
    int n = x.size();
    double y = 0;

    for (int i = 0; i < n; ++i) {
        y += x[i] / n;
    }
    return y;
}')
```
